hormones

### 2019-10-11 analysis/07_hormones.Rmd
three sets of hormone figs

```{r clean}
charlevels <- c("control", "bldg", "lay", "inc_d3", "inc_d9", "inc_d17", "hatch", "n5", "n9" )
removelevels <- c("M_inc3",  "M_inc9", "M_inc17", "M_n2")
maniplevels <- c("early", "prolong", "extend")
combolevels <- c("control", "bldg", "lay", "inc_d3", "M_inc3",
                  "inc_d9",  "M_inc9", "early",  "inc_d17","M_inc17",  "hatch",
                 "prolong", "extend", "M_n2",
                  "n5", "n9" )
```

@@

PETC <- PETC %>%
    mutate(sex = fct_recode(sex,
                            "female" = "f",
                            "male" = "m"),
           study = fct_collapse(stage,
                                 characterization = charlevels,
                                manipulation = maniplevels,
                                removal = removelevels)) %>% 
  drop_na()
colnames(PETC)[colnames(PETC)=="stage"] <- "treatment"
PETC <- PETC %>% select(study, treatment, sex, hormone, plasma_conc)
summary(PETC)
```

@@

hormones$okay <- ifelse(hormones$hormone == "cort" & hormones$plasma_conc > 30, "bad",
                    ifelse(hormones$hormone == "progesterone" & hormones$plasma_conc > 5, "bad", 
                           ifelse(hormones$hormone == "prolactin" & hormones$plasma_conc > 150, "bad", 
                            ifelse(hormones$hormone == "prolactin" & hormones$treatment == "control", "bad", 
                        ifelse(hormones$hormone == "testosterone" & hormones$sex == "female", "bad",
                               ifelse(hormones$hormone == "estradiol" & hormones$sex == "male", "bad", "okay"))))))
hormones <- hormones %>% filter(okay == "okay") %>% droplevels()

prl.char <- hormones %>% filter(hormone == "prolactin", study == "characterization")   %>%  droplevels()
test.char <- hormones %>% filter(hormone == "testosterone", treatment %in% charlevels)   %>%  droplevels()
est.char <- hormones %>% filter(hormone == "estradiol", treatment %in% charlevels)   %>%  droplevels()
prog.char <- hormones %>% filter(hormone == "progesterone", treatment %in% charlevels)   %>%  droplevels()
cort.char <- hormones %>% filter(hormone == "cort", treatment %in% charlevels)   %>%  droplevels()

@@

```{r characterization-hormones}

hormonecharplot <- function(myhormone, myylab){

@@

```{r aov}

aovSexTretment <- function(mydata, whichormone){
  aov2 <- aov(data = mydata, plasma_conc ~ treatment + sex)
  print(whichormone)
  print(summary(aov2))
  print(TukeyHSD(aov2, which = "treatment"))
}
aovSexTretment(prl.char, "PRL")
aovSexTretment(cort.char, "CORT")
aovSexTretment(prog.char, "PROG")

aovTretment  <- function(mydata, whichormone){
  aov1 <- aov(data = mydata, plasma_conc ~ treatment )
  print(whichormone)
  print(summary(aov1))
  print(TukeyHSD(aov1, which = "treatment"))
}

@@

>>>>

```{r clean}
charlevels <- c("control", "bldg", "lay", "inc_d3", "inc_d9", "inc_d17", "hatch", "n5", "n9" )
removelevels <- c("M_inc3",  "M_inc9", "M_inc17", "M_n2")
maniplevels <- c("early", "prolong", "extend")
combolevels <- c("control", "bldg", "lay", "inc_d3", "M_inc3",
                  "inc_d9",  "M_inc9", "early",  "inc_d17","M_inc17",  "hatch",
                 "prolong", "extend", "M_n2",
                  "n5", "n9" )

manipchar <- c("early",  "inc_d17", "hatch", "prolong", "extend", "n5")
removechar <- c("inc_d3", "M_inc3",  "inc_d9", "M_inc9","inc_d17", "M_inc17","hatch", "M_n2")
```

@@

PETC <- PETC %>%
    mutate(sex = fct_recode(sex,
                            "female" = "f",
                            "male" = "m"),
           study = fct_collapse(stage,
                                characterization = charlevels,
                                manipulation = maniplevels)) %>% 
  drop_na()
colnames(PETC)[colnames(PETC)=="stage"] <- "treatment"
PETC <- PETC %>% select(study, treatment, sex, hormone, plasma_conc)
summary(PETC)
```

@@

hormones$okay <- ifelse(hormones$hormone == "cort" & hormones$plasma_conc > 30, "bad",
                    ifelse(hormones$hormone == "progesterone" & hormones$plasma_conc > 5, "bad", 
                           ifelse(hormones$hormone == "prolactin" & hormones$plasma_conc > 150, "bad", 
                            ifelse(hormones$hormone == "prolactin" & hormones$treatment == "control", "bad", 
                        ifelse(hormones$hormone == "testosterone" & hormones$sex == "female", "bad",
                               ifelse(hormones$hormone == "estradiol" & hormones$sex == "male", "bad", "okay"))))))
hormones <- hormones %>% filter(okay == "okay") %>% droplevels()
summary(hormones)

prl.char <- hormones %>% filter(hormone == "prolactin", treatment %in% charlevels)   %>%  droplevels()
test.char <- hormones %>% filter(hormone == "testosterone", treatment %in% charlevels)   %>%  droplevels()
est.char <- hormones %>% filter(hormone == "estradiol", treatment %in% charlevels)   %>%  droplevels()
prog.char <- hormones %>% filter(hormone == "progesterone", treatment %in% charlevels)   %>%  droplevels()
cort.char <- hormones %>% filter(hormone == "cort", treatment %in% charlevels)   %>%  droplevels()

@@

```{r characterization}
hormonecharplot <- function(myhormone, myylab){

@@

```{r aov-char}
aovSexTretment <- function(mydata, whichormone){
  aov2 <- aov(data = mydata, plasma_conc ~ treatment + sex)
  print(whichormone)
  print(summary(aov2))
  #print(TukeyHSD(aov2, which = "treatment"))
}

aovTretment  <- function(mydata, whichormone){
  aov1 <- aov(data = mydata, plasma_conc ~ treatment )
  print(whichormone)
  print(summary(aov1))
  #print(TukeyHSD(aov1, which = "treatment"))
}

aovSexTretment(prl.char, "PRL")
aovSexTretment(cort.char, "CORT")
aovSexTretment(prog.char, "PROG")

@@

```{r manipulation}
hormonemanipPRL <- function(myhormone, myylab){

  hormones %>% 
    filter(treatment %in% manipchar,
           hormone %in% c(myhormone))  %>% 
  ggplot(aes(x = treatment, y = plasma_conc, fill = sex)) +
    geom_boxplot() + 
    theme(axis.text.x = element_text(angle = 45, hjust = 1),
          legend.position = "none",
          strip.text = element_blank()) +
    scale_fill_manual(values = mycolors) +
    labs(y = myylab, x = NULL) +
    facet_wrap(~sex) +
    annotate("rect", xmin = 0.5, xmax = 1.5, ymin = 0, ymax = 125, alpha = .2) +
    annotate("rect", xmin = 3.5, xmax = 5.5, ymin = 0, ymax = 125, alpha = .2) 
}

hormonemanipPRL("prolactin", "PRL (ng/mL)")

hormonemanipSteroids <- function(myhormone, myylab, myymax){

  hormones %>% 
    filter(treatment %in% manipchar,
           hormone %in% c(myhormone))  %>% 
  ggplot(aes(x = treatment, y = plasma_conc, fill = sex)) +
    geom_boxplot() + 
    theme(axis.text.x = element_text(angle = 45, hjust = 1),
          legend.position = "none",
          strip.text = element_blank()) +
    scale_fill_manual(values = mycolors) +
    labs(y = myylab, x = NULL) +
    facet_wrap(~sex) +
    annotate("rect", xmin = 2.5, xmax = 4.5, ymin = 0, ymax = myymax, alpha = .2)
}

a <- hormonemanipSteroids("estradiol", "E (ng/mL)", 1)
b <- hormonemanipSteroids("testosterone", "T (ng/mL)", 3.5)
c <- hormonemanipSteroids("cort", "CORT (ng/mL)", 10)
d <- hormonemanipSteroids("progesterone", "PROG (ng/mL)",2.5)

plot_grid(d,a,c,b, rel_widths = c(0.65,0.35))
```
```{r removal}
hormoneremoval <- function(myhormone, myylab){

  hormones %>% 
    filter(treatment %in% removechar,
           hormone %in% c(myhormone))  %>% 
  ggplot(aes(x = treatment, y = plasma_conc, fill = sex)) +
    geom_boxplot() + 
    theme(axis.text.x = element_text(angle = 45, hjust = 1),
          legend.position = "none",
          strip.text = element_blank()) +
    scale_fill_manual(values = mycolors) +
    labs(y = myylab, x = NULL) +
    facet_wrap(~sex) +
    annotate("rect", xmin = 1.5, xmax = 2.5, ymin = 0, ymax = 125, alpha = .2) +
    annotate("rect", xmin = 3.5, xmax = 4.5, ymin = 0, ymax = 125, alpha = .2) +
    annotate("rect", xmin = 5.5, xmax = 6.5, ymin = 0, ymax = 125, alpha = .2) + 
    annotate("rect", xmin = 7.5, xmax = 8.5, ymin = 0, ymax = 125, alpha = .2) 
}

hormoneremoval("prolactin", "PRL (ng/mL)")
```
